final-project

### 2024-11-25 code/final-project.Rmd
tables

## agecat
CrossTable(agecat, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## race
CrossTable(racecat, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

@@

## grade
CrossTable(grade, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE, useNA = "always")

@@

## agecat
survfit(Surv(survival_months, survival_status)~agecat)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ agecat), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ agecat)

## race
survfit(Surv(survival_months, survival_status)~racecat)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ racecat), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ racecat)

@@

# run the cox ph model
# this will be in Table 2 as multivariate analysis
model <- coxph(
  Surv(survival_months, survival_status) ~ 
    relevel(factor(subtype), ref = '0') +
    relevel(factor(agecat), ref = '0') + 
    relevel(factor(racecat), ref = '0') +
    relevel(factor(marriage_status),ref='1') + 
    relevel(factor(n_stage),ref='0') + 
    relevel(factor(m_stage),ref='0') +
    relevel(factor(grade), ref='0') +
    relevel(factor(radiation), ref='1') + 
    relevel(factor(chemotherapy), ref='1') +
    relevel(factor(surgery), ref='1'),
  data = seer, 
  method = "breslow")

@@

# get the correlations between residuals and ranked survival times
apply(resid$y, 2, \(x) cor.test(rank(resid$time), x))
# subtype1: -0.034
# subtype2: -0.1296
# subtype3: -0.0435
# agecat1: -0.0331
# agecat2: -0.0304
# racecat1: 0.0077
# racecat2: 0.0102
# marraige_status0: 0.0043
# n_stage1: 0.0188
# m_stage: -0.0046
# grade: -0.0412
# radiation: -0.072
# chemotherapy: -0.1122
# surgery: -0.0715

@@

# Now, we can correctly mutate and format the output
cor_df <- cor_df_all %>%
  mutate(
    Correlation = paste(correlation.cor, " (", conf_low, ", ", conf_high, ")", sep = ""),
    P_Value = as.character(p_value)
  ) %>%
  select(Correlation, P_Value)
rownames(cor_df) <- c(
  "subtype1",
  "subtype2",
  "subtype3",
  "agecat1",
  "agecat2",
  "racecat1",
  "racecat2",
  "marraige_status0",
  "n_stage1",
  "m_stage",
  "grade",
  "radiation",
  "chemotherapy",
  "surgery"
)

@@

options(scipen=999)
model_ext <- coxph(
  formula = Surv(
    seer.cp$start, seer.cp$survival_months, seer.cp$survival_status) ~ 
    subtype1 + subtype2 + subtype3 + 
    relevel(factor(agecat), ref = '0') + 
    relevel(factor(racecat), ref = '0') +
    relevel(factor(marriage_status),ref='1') + 
    relevel(factor(n_stage),ref='0') + 
    relevel(factor(m_stage),ref='0') +
    relevel(factor(grade), ref='0') +
    relevel(factor(radiation), ref='1') + 
    relevel(factor(chemotherapy), ref='1') +
    relevel(factor(surgery), ref='1') +
    subtype1_time + subtype2_time + subtype3_time, 
  data = seer.cp, method="breslow")

>>>>

## age
CrossTable(age, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## race
CrossTable(race, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

@@

## grade
CrossTable(grade, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

@@

## age
survfit(Surv(survival_months, survival_status)~age)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ age), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ age)

## race
survfit(Surv(survival_months, survival_status)~race)
summary(survfit(Surv(time = survival_months, event = survival_status) ~ race), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ race)

@@

# run the cox ph model
# this will be in Table 2 as multivariate analysis
model <- coxph(
  Surv(survival_months, survival_status) ~ 
    relevel(factor(subtype), ref = '0') +
    relevel(factor(age), ref = '0') + 
    relevel(factor(race), ref = '0') +
    relevel(factor(marriage_status),ref='1') + 
    relevel(factor(n_stage),ref='0') + 
    relevel(factor(m_stage),ref='0') +
    relevel(factor(grade), ref='0') +
    relevel(factor(radiation), ref='1') + 
    relevel(factor(chemotherapy), ref='1') +
    relevel(factor(surgery), ref='1'),
  data = seer, 
  method = "breslow")

@@

# get the correlations between residuals and ranked survival times
apply(resid$y, 2, \(x) cor.test(rank(resid$time), x))
# subtype1: -0.034
# subtype2: -0.1296
# subtype3: -0.0435
# age1: -0.0331
# age2: -0.0304
# race1: 0.0077
# race2: 0.0102
# marraige_status0: 0.0043
# n_stage1: 0.0188
# m_stage: -0.0046
# grade: -0.0412
# radiation: -0.072
# chemotherapy: -0.1122
# surgery: -0.0715

@@

# Now, we can correctly mutate and format the output
cor_df <- cor_df_all %>%
  mutate(
    Correlation = paste(correlation.cor, " (", conf_low, ", ", conf_high, ")", sep = ""),
    P_Value = as.character(p_value)
  ) %>%
  select(Correlation, P_Value)
rownames(cor_df) <- c(
  "subtype1",
  "subtype2",
  "subtype3",
  "age1",
  "age2",
  "race1",
  "race2",
  "marraige_status0",
  "n_stage1",
  "n_stage2",
  "n_stage3",
  "m_stage",
  "grade",
  "radiation",
  "chemotherapy",
  "surgery"
)

@@

options(scipen=999)
model_ext <- coxph(
  formula = Surv(
    seer.cp$start, seer.cp$survival_months, seer.cp$survival_status) ~ 
    subtype1 + subtype2 + subtype3 + 
    relevel(factor(age), ref = '0') + 
    relevel(factor(race), ref = '0') +
    relevel(factor(marriage_status),ref='1') + 
    relevel(factor(n_stage),ref='0') + 
    relevel(factor(m_stage),ref='0') +
    relevel(factor(grade), ref='0') +
    relevel(factor(radiation), ref='1') + 
    relevel(factor(chemotherapy), ref='1') +
    relevel(factor(surgery), ref='1') +
    subtype1_time + subtype2_time + subtype3_time, 
  data = seer.cp, method="breslow")
